### 2018-01-08 response_to_referee.txt
start referee response

Dear Frank:

Please find enclosed our revised version of MS#AAS07769, Data analysis
recipes: Using Markov Chain Monte Carlo by Hogg & Foreman-Mackey.

In what follows, the referee comments are indented and our responses
are not.

Thank you,

David

     ---------------------------------------------------------------------- 
     Referee Report

     The authors provide a simple, and somehow anecdotal discussion
     about MCMC methods (for astronomers, although the title don't
     explicitly says it) in an informal and pragmatic style.

We figured that because this is in the AAS journals, that would supply
the relevant context, though we wouldn't object to modifying the title
slightly.

     The concept of conveying an intuitive description of statistical
     methods for the broad astronomical community is certainly
     something I support, however I would like to ask the authors and
     editor to consider some caveats below.

     Novelty and references

     The draft lacks of some important literature contextualization. A
     critical one is a recent Annual Review of Astronomy and
     Astrophysics entitled Markov Chain Monte Carlo Methods for
     Bayesian Data Analysis in Astronomy
     https://arxiv.org/abs/1706.01629

Yes, we know this paper. We now cite it [WHERE].

     I would like the authors to please explain how their review
     brings something new to the literature, considering the amount of
     good and available material about MCMC.

Good point. We are more explicit about this [WHERE]. The idea is to
delve deeper here, but on MCMC specifically, but not so much about new
developments; more like we are trying to specifically emphasize the
importance of basic good pratices; we say that more clearly now.

     Such as e.g.

     Doing Bayesian Data Analysis, Second Edition: A Tutorial with R,
     JAGS, and Stan, by John Kruschke, and many others

We also now cite some of this literature [WHERE; DO THIS].

     A potential reason is to update the readers about recent
     developments in the field.  But again, the authors seems to
     ignore the current astronomical literature. For example, the
     authors are aware of Statistical Bayesian packages such as JAGS
     and Stan, but do not cite two recent books in Astronomy that make
     use of them, such as:

     Bayesian Methods for the Physical Sciences: Learning from
     Examples in Astronomy and Physics, Springer Series in
     Astrostatistics, 2015 Andreon, and Weaver

     And

     Bayesian Models for Astrophysical Data: Using R, JAGS, Python,
     and Stan, Cambridge University press, Hilbe, de Souza and Ishida,
     2017

Good point; these are all now cited [WHERE].

     2. Confidence versus credible intervals

     I am sure the authors are aware of the difference between both,
     and should know they don't even agree beyond simple cases: such
     as for normally distributed data. Hence the quote "Because they
     are so afraid of being confused with frequentists, Bayesians
     often call these regions "credible" rather than "confidence""

     Not only seems inappropriate for a scientific paper, but it is
     actually misleading. See for instance The fallacy of placing
     confidence in confidence intervals, Morey et al. 2015
     https://www.ncbi.nlm.nih.gov/pmc/articles/PMC4742505/

     Moreover, such discussion -- Bayesian vs Frequentist -- has far
     deeper roots. See e.g. Probability theory, the logic of science,
     E.T. Jaynes, Cambridge University Press 2003.

Actually, our problem is with the *names*. We agree with the referee
that these two things are importantly and substantially different! We
have fixed that all up in the text now [DO THIS].

     3. Lack of discussion about prediction vs credible intervals.

[UHHH SUP?]

     4. Model check

     Lack of contemporary discussion such as e.g. Gabry et
     al. Visualization in Bayesian workflow,
     https://arxiv.org/abs/1709.01449

Ooh that's a great reference! We cite this and were pleased to be
pointed at this.

     5. MORE SOPHISTICATED SAMPLING METHODS

     Please consider adding some material about

     Approximate Bayesian Computation, since it has gained space in astronomy e.g.

     https://arxiv.org/abs/1608.07606
     https://arxiv.org/abs/1607.01782 
     https://arxiv.org/abs/1504.06129 
     https://arxiv.org/abs/1202.1426

Good point -- we are even working on this ourselves. We added a small
bit of discussion about this [WHERE].

     ---------------------------------------------------------------------- 
     Statistics Editor Comments:

     As a didactic presentation, I suggest two additions:

     (a) Add a table listing MCMC approaches so the reader appreciates
     the variety of options available
     (e.g. https://www.rdocumentation.org/packages/LaplacesDemon/versions/16.0.1/topics/LaplacesDemon)

     (b) Discuss how to choose when MCMC sampling and when
     optimization (e.g. EM Algorithm or simplex) is appropriate. The
     former is commonly used for characterizing Bayesian posteriors
     while the latter are commonly used for maximizing likelihoods.